Relative Stable Maps in Genus One via Central Alignments

### 2018-11-13 new.tex
a new plan of conquest

\newcommand{\pM}{\pazocal{M}}
\newcommand{\TT}{\operatorname{T}}
\newcommand{\oM}{\overline{\mathcal{M}}}
\newcommand{\M}[4]{\overline{\mathcal{M}}_{#1,#2}(#3,#4)}
\newcommand{\Q}[4]{\mathcal{Q}_{#1,#2}(#3,#4)}
\newcommand{\Qe}[4]{\mathcal{Q}^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\Qt}[4]{\widetilde{\mathcal Q}_{#1,#2}(#3,#4)}
\newcommand{\QG}[4]{\mathcal{Q}G_{#1,#2}(#3,#4)}
\newcommand{\QGe}[4]{\mathcal{Q}G^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\D}[3]{\mathcal{D^Q}(#1,#2,#3)}
\newcommand{\E}[3]{\mathcal{E^Q}(#1,#2,#3)}
\newcommand{\PP}{\mathbb P}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\tVZc}[4]{\widetilde{\mathcal{V\!Z}}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZc}[4]{\mathcal{V\!Z}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZcLi}[4]{\mathcal{V\!Z}^{\rm{ctr, Li}}_{#1,#2}(#3,#4)}
\newcommand{\VZrel}[4]{\mathcal{V\!Z}^{\rm{rel}}_{#1,#2}(#3,#4)}
\newcommand{\stab}{\rm{stab}}
\newcommand{\stC}{C'}
\newcommand{\stf}{f'}
\newcommand{\stpi}{\pi'}
\newcommand{\sts}{s'}
\newcommand{\N}{\mathbb{N}}
\newcommand{\OO}{\mathcal{O}}
\renewcommand{\to}{\rightarrow}
\newcommand{\A}{\mathcal A}
\newcommand{\B}{\mathcal B}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\EE}{\mathbf{E}}
\renewcommand{\L}{\mathcal L}
\newcommand{\LL}{\mathbf{L}}
\newcommand{\MM}{\mathfrak M}
\newcommand{\Aaff}{\mathbb{A}}
\newcommand{\kfield}{\Bbbk}
\newcommand{\comp}{\chi}
\newcommand{\sst}{\sigma^{\operatorname{ss}}}
\newcommand{\Pic}{\operatorname{Pic}}
\newcommand{\Def}{\operatorname{Def}}
\newcommand{\Spec}{\operatorname{Spec}}
\newcommand{\Proj}{\operatorname{Proj}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Gm}{\mathbb{G}_{\text{m}}}
\newcommand{\virt}[1]{[#1]^{\operatorname{virt}}}
\newcommand{\vip}[1]{[#1]^{\operatorname{prod}}}
\newcommand{\Id}{\operatorname{Id}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\HH}{\operatorname{H}}
\newcommand{\Achow}{\operatorname{A}}
\newcommand{\pt}{\operatorname{pt}}
\newcommand{\bq}{\begin{equation}}
\newcommand{\eq}{\end{equation}}
\newcommand{\ba}{\begin{aligned}}
\newcommand{\ea}{\end{aligned}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bsm}{\left(\begin{smallmatrix}}
\newcommand{\esm}{\end{smallmatrix}\right)}                   
\newcommand{\bpm}{\begin{pmatrix}}
\newcommand{\epm}{\end{pmatrix}}
\newcommand{\barr}{\begin{displaymath}\begin{array}{cccc}}
\newcommand{\earr}{\end{array}\end{displaymath}}
\newcommand{\barrl}{\begin{displaymath}\begin{array}{lcl}}
\newcommand{\earrl}{\end{array}\end{displaymath}}
\newcommand{\barl}{\begin{displaymath}\begin{array}{l}}
\newcommand{\earl}{\end{array}\end{displaymath}}
\newcommand{\bxym}{ \begin{displaymath}\xymatrix }
\newcommand{\exym}{\end{displaymath}}
\newcommand{\bcd}{\begin{center}\begin{tikzcd}}
\newcommand{\ecd}{\end{tikzcd}\end{center}}
\newcommand{\R}{\operatorname{R}^{\bullet}}
\newcommand{\dvr}{\Delta}
%\newcommand{\sslash}{\mathbin{/\mkern-6mu/}}
\newcommand{\tr}{{\rm tr}}
\newcommand{\Isom}{\text{Isom}}
\newcommand{\pr}{\operatorname{pr}}
\newcommand{\ev}{\operatorname{ev}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\vdim}{\operatorname{vdim}}
\newcommand{\ildef}[1]{\emph{#1}}
\newcommand{\om}[1]{\mathcal{#1}}
\newcommand{\h}{\operatorname{h}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\RR}{\textbf{R}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\ovm}[1]{\overline{\mathcal{#1}}}
\newcommand{\ovt}[1]{\widetilde{\mathcal{#1}}}
\newcommand{\ov}[1]{\overline{#1}}

@@

\section{A desingularisation of Chen-Kim's moduli space of genus one logarithmic maps to expanded pairs, for $(\PP^N|H)$}

\subsection{Relative logarithmic maps to expanded pairs} We recall a number of definitions from \cite{ChenDegeneration}; see also \cite{KimLog}.

Fix a smooth pair $(X,Y)$, i.e. a smooth projective variety $X$ with a smooth divisor $Y$.

\begin{definition}
 An \emph{extended log twisted expansion} of $(X,Y)$ is a quadruple $(\pi\colon W\to S, \pi_X\colon W\to X, D\subseteq W, \pM_S^{W/S}\to\pM_S)$ such that:
 \begin{enumerate}
  \item $\pi$ makes $W$ into a flat, proper algebraic space over $S$, with fibers at worst nodal in codimension 1, admitting a lifting to a special log morphism $\pi\colon (W,\pM_W^{W/S})\to(S,\pM_S^{W/S})$, such that:
  \begin{enumerate}
   \item $\pM_W^{W/S}$ and $\pM_S^{W/S}$ are locally free log structures;
   \item for every geometric point $\bar{w}\to W$, either $\bar{w}$ is in the smooth locus of $\pi_S$ and $\pi^\flat\colon \pi^*\overline{\pM}_{S,\bar{w}}^{W/S}\to \overline{\pM}_{W,\bar{w}}^{W/S}$ is an isomorphism, or $\bar{w}$ is in the nodal locus and the following diagram is cocartesian:
   \bcd
   \NN\ar[r,"{(1,1)}"]\ar[d,"h"] & \NN^2\ar[d] \\
   \pi^*\overline{\pM}_{S,\bar{w}}^{W/S}\ar[r] & \overline{\pM}_{W,\bar{w}}^{W/S}
   \ecd
   with $h(e)$ irreducible;
   \item for every geometric point $\bar{s}\to S$ there is a bijection between the irreducible elements of $\overline{\pM}_{S,\bar{s}}^{W/S}$ and the irreducible components of $W_{\bar{s}}^{\text{sing}}$.
  \end{enumerate}
  \item $D\subseteq W$ is a Cartier divisor, smooth over $S$, and such that $\pi_X(D)=Y$;
 \item $\pM_S$ is twisted by integers $(r_1,\ldots,r_l)$ and expanded with respect to $\pM_S^{W/S}$, namely \'{e}tale locally at every point $s\in S$ the morphism of log structures $\pM_S^{W/S}\to\pM_S$ admits a chart:
 \bcd
 \NN^l\ar[d]\ar[r,"{(r_1,\cdots,r_l)}"] &\NN^l\ar[r,"{(id,0)}"]&\NN^l\oplus\NN^k\ar[d] \\
 \pM_S^{W/S}\ar[rr] & & \pM_S
 \ecd

 \end{enumerate}
\end{definition}
 The log structure on $W$ is then given by
 \[ \pM_W=\pi^*\pM_S\oplus_{\pi^*\pM_S^{W/S}}\pM_W^{W/S}\oplus_{\OO^*_W}\pM^D\]
 where $\pM_D$ is the divisorial log structure associated to $D$. 
Notice that with this definition $(W,\pM_W)\to(S,\pM_S)$ is log smooth.

Similarly a \emph{log twisted curve} is $((C,p_1,\ldots,p_n)\to S, \pM_S^{C/S}\to\pM_S)$ with
\[\pM_C=\pi^*\pM_S\oplus_{\pi^*\pM_S^{C/S}}\pM_C^{C/S}\oplus_{\OO^*_C}\bigoplus_{i=1}^n\pM^{p_i}\]
A log twisted curve is \emph{minimal} if the log structure $\pM_S$ is locally free, and for every geometric point $\bar{s}\to S$ and irreducible element $b\in\overline{\pM}_S$, we can find an irreducible $a\in\overline{\pM}^{C/S}_S$ and a positive integer $l$ such that $a\mapsto lb$ (by definition this means that the morphism of log structures $\pM_S^{C/S}\to\pM_S$ is \emph{simple}). So essentially we are allowed to take roots of the smoothing parameters of the nodes of $C_{\bar{s}}$.

Finally, fix an $n$-tuple of non-negative contact orders $(c_1,\ldots,c_n)$.
\begin{definition}
 A \emph{relative log prestable map} with tangency condition $\mathbf{c}$ is given by a diagram of log morphisms:
 \bcd
 ((C,\mathbf p),\pM_C)\ar[rr,"f"]\ar[dr] & & ((W,D),\pM_W)\ar[rr,"\pi_X"]\ar[dl] & & (X,Y)\times S \\
  & (S,\pM_S) & & &
 \ecd
 \begin{enumerate}
  \item $(C,\mathbf p),\pM_C)\to (S,\pM_S)$ is a minimal log twisted curve;
  \item $((W,\pM_W)\to (S,\pM_S),(W,D)\to(X,Y))$ is an extended twisted log expansion of $(X,Y)$;
  \item (\emph{corank condition}) \sidenote{split the definition into a more general one over LogSch and minimality?} for every geometric point $\bar{s}\in S$, the rank of \[\operatorname{Coker}(\pM^{W/S}_{S,\bar{s}}\to \pM_{S,\bar{s}})\] is equal to the number of non-distinguished nodes of $C_{\bar{s}}$; recall that a node of $C_{\bar{s}}$ is \emph{distinguished} if it is mapped under $f$ to the singular locus of $W_{\bar{s}}$.
  \item (\emph{log admissibility}) the morphism of log structures $f^\flat\colon f^*\pM_W\to\pM_C$ is simple at every distinguished node;
  \item (\emph{tangency condition}) the underlying map $\underline{f}\colon C\to (W,D)$ is non-degenerate, namely no component of $C$ is mapped entirely into $D$, and locally around every marking $p_i$ we can find a chart
  \bcd
  \NN\ar[r,"\cdot c_i"]\ar[d] & \NN\ar[d] \\
  f^*\pM^D\ar[r] & \pM^{p_i}
  \ecd
  for the restriction of $f^\flat$ to the $f^*\pM^D$ factor.
  \end{enumerate}
\end{definition}
\begin{definition}
 A relative log prestable map as above is \emph{stable} if for every geometric point $\bar{s}\in S$ the group of automorphisms $(\sigma,\tau)$ is finite, where:
 \begin{enumerate}
  \item $\sigma$ is an automorphism of $((C,\pM_C)\to (S,\pM_S))_{\bar{s}}$ preserving the markings $\mathbf p$;
  \item $\tau$ is an automorphism of $((W,\pM_W)\to (S,\pM_S))_{\bar{s}}$ preserving $D$ and $\pi_X\colon W_{\bar{s}}\to X$;
  \item $\tau\circ f_{\bar{s}}=f_{\bar{s}}\circ \sigma$.
 \end{enumerate}
\end{definition}
\begin{remark}
 The corank and simplicity condition imply the following: the rank of the locally free log structure $\pM_S$ at $\bar{s}$ amounts to the number of irreducible components of $W_{\bar{s}}^{\text{sing}}$ plus the number of non-distinguished nodes of $C_{\bar{s}}$; all the smoothing parameters of the distinguished nodes mapping to the same component of $W_{\bar{s}}^{\text{sing}}$ have been identified \emph{up to twisting}. In facts, at a distinguished node $q$ of $C$, letting $R$ be the local ring of $S$ at $\pi(q)$, we can find \'{e}tale local charts
 \bcd
 \NN^2\oplus_{\NN} \pM_{S}\ar[r,"{(id,\cdot m_q)}" below]\ar[d] & \NN^2\oplus_{\NN} \pM_{S}\ar[d] \\
 f^*\pM_W \ar[r,"f^\flat" below]\ar[d] & \pM_C\ar[d] \\
 R[z_1,\ldots,z_{\dim(X)-1},x,y]/(xy-t)\ar[r,"\left(\substack{x\mapsto a^{m_q}\\ y\mapsto b^{m_q}\\ t\mapsto s^{m_q}}\right)" below=.2cm] & R[a,b]/(ab-s)
 \ecd
 where the twist at $f(q)$ is given by $l$, the one at $q$ is given by $r_q$, and the equality $l=r_qm_q$ has to hold.
\end{remark}

\subsection{The desingularisation via central alignments} The following is a variation on \cite[\S 3.3]{RSPW2}.

\section{Gathmann's recursive structure on the boundary of maps to expanded pairs}

We find a natural line bundle with a section whose vanishing locus consists of those maps where the target splits and the marking $p_k$ is mapped to a higher level of the accordion. We express these boundary strata recursively as fibered products of desingularised relative and rubber spaces with lower numerical invariants; the multiplicity calcultion is performed for $\PP^1|\{\infty\}$ and rests on work of Vakil \cite{Vre}. Putting everything together we reobtain Gathmann's algorithm for computing relative invariants, or the invariants of a smooth hyperplane section.

@@

\begin{remark}
 Notice that gluing does not present the subtleties to which log geometry has made us used, since we endow the curves with the minimal centrally aligned structure, possibly after prescribing that the smoothing parameters for the newly created nodes are \emph{bigger} than the contraction radius $\delta$.
\end{remark}

\begin{remark}
 A dimensional computation shows that $\mathcal Y^b$ is never relevant to the purpose of the recursion formula. On the other hand the classes of $\mathcal Y^a$ and $\mathcal Y^c$ should be weighted with the coefficient $\frac{\prod m^{(i)}}{r!}$. Let $D_{1,\alpha,k}(\PP^N|H)$ be the union of all the $\mathcal Y^a$ and $\mathcal Y^c$ as above, for any choice of number of external components $r$, distribution of the markings $A$, and of the degree $D$, provided the marked point $x_k$ lies on the internal component $C^{(0)}$. Let each component be endowed with its fundamental class weighted as above and let $\virt{D_{1,\alpha,k}(\PP^N|H)}$ be the sum of all these classes.
\end{remark}

\begin{thm}
 Assume $\sum\alpha=d$. Then \[(\alpha_k\psi_k+\ev_k^*H)[\VZc{1}{\alpha}{\PP^N|H}{d}]=\virt{D_{1,\alpha,k}(\PP^N|H)}.\]
\end{thm}
\begin{proof}
We perform a number of reductions.
 \begin{itemize}%[leftmargin=1cm]
  \item \textbf{Reduction to the case of $\PP^1|\{\infty\}$:} by choosing a general $N-2$-plane $A\subseteq H$ and projecting away from it. Must show that the rational map $\rho_A\colon \VZc{1}{n}{\PP^N|H}{d}\dashrightarrow \VZc{1}{n}{\PP^1|\{\infty\}}{d}$ is defined and smooth at a general point of $D_{1,\alpha,k}(\PP^N|H)$.

  \item \textbf{Reduction to an \'etale neighbourhood of $C^{(0)}$:} the plan is to understand the global perfect obstruction theory of $\VZc{1}{n}{\PP^1}{d}$ and see that the tangent to the latter is a sum of contributions from the \emph{special loci}, namely contracted components, ramification points, markings and nodes.

  \begin{remark}
   When we are studying the infinitesimal deformation theory of centrally aligned curves, notice that being centrally aligned must be checked on geometric points, and minimality is open, hence any deformation as a log smooth curve will automatically be minimal centrally aligned. The tangent space is then $H^1(C,T_C^{\rm{log}})$; notice that this is independent from the log structure on the base.
  \end{remark}

  \begin{example}
   There are non-trivial deformations that are trivial on the underlying marked prestable curve. If $(C,\mathcal M_C)\to (\Spec(k),Q^{\rm{cen}})$ is a centrally aligned curve, then in order to define a deformation over $\Spec(k[\epsilon])$ we need to extend the log structure on the base by choosing a map $Q^{\rm{cen}}\to k[\epsilon]$. Notice that if we want the underlying curve to be a trivial deformation we better send all the smoothing parameters to $0$ in $k[\epsilon]$; on the other hand, if $Q^{\rm{cen}}$ contains an element such as $e_i-e_j$ (making $R_i$ further from the circuit than $R_j$), then we are free to map it to an invertible element in $k[\epsilon]$, which will kill it in the associated log structure;\sidenote{Invertible elements of $k[\epsilon]$ remain invertible when setting $\epsilon=0$, which seems to kill $e_i-e_j$ on the central fiber too.} this will make $R_i$ and $R_j$ be \emph{at the same distance} from the circuit in the deformation. Importantly, $Q^{\rm{log}}\to Q^{\rm{cen}}$ is an epimorphism only in the category of integral monoid (of which the multiplicative monoid of a ring is never an object, because it has a sink).
  \end{example}

 \end{itemize}

\end{proof}

\appendix
\section{Old}

\subsection{Definition of the relative space} We now give a definition of the relative space inside the moduli space of centrally aligned stable maps.
\begin{dfn}
Let $X$ be a smooth projective variety with a smooth divisor $Y$. Fix a vector of tangency conditions $\alpha=(\alpha_1,\ldots,\alpha_n)$ with $\Sigma_{i=1}^n \alpha_i = Y \cdot \beta$. Then the \emph{centrally aligned relative space} 
\begin{equation*}\VZc{1}{\alpha}{X|Y}{\beta} \end{equation*}
is defined to be the closed substack of $\VZc{1}{n}{X}{\beta}$ satisfying the following conditions:
\begin{enumerate}
\item \emph{Gathmann's relative condition:} for each connected component $Z$ of $f^{-1}(Y)$, either:
\begin{enumerate}
\item $Z$ is a single point of $C$, in which case if $Z=x_i$ is a marking then we require that the multiplicity of $f$ at $x_i$ along $Y$ is at least $\alpha_i$ (if $Z$ is not a marking then there is no condition);
\item $Z$ is a subcurve $Z \subseteq C$, in which case we require that
\begin{equation*} f^*[Y]|_Z -\sum_{x_i\in Z}\alpha_ix_i\end{equation*}
is an effective class in $\Achow_0(Z)$.
\end{enumerate}

\item \emph{Compatibility condition:} this applies when the minimal subcurve of genus one (the \emph{circuit} or \emph{core}) is contracted to a point in $Y$. In this case we require that the following diagram of solid arrows may be completed to a commutative diagram:
\bcd
Q_{\rm{min}}^{\rm{prestable}}\ar[r]\ar[d] & Q_{\rm{min}}^{\rm{log.map}} \\
Q_{\rm{min}}^{\rm{cen.align}}\ar[dashed,ur]
\ecd
\end{enumerate}
\end{dfn}

\begin{remark}[Explaining Gathmann]
Gathmann's condition is promptly extended to the case that $\Sigma_{i=1}^n \alpha_i \leq Y \cdot \beta$. \textcolor{red}{Question: what is the correct generalisation of the compatibility condition?} Here we try to make the condition for a curve component of $f^{-1}(Y)$ into an easily verifiable criterion: if $Z$ is a smooth elliptic curve, recall that every line bundle of positive degree on it is effective, hence the relative condition can be rephrased as the following numerical criterion:
\begin{equation*} f_*[Z]\cdot [Y]+\sum_{j=1}^r m^{(j)}\geq \sum_{x_i\in Z}\alpha_i \end{equation*}
together with the additional requirement that, when the above is an equality, there is an isomorphism of line bundles:
\begin{equation}\label{eqn:eq_in_Pic}
(f|_{Z})^*\OO_X(Y) \otimes \OO_Z\left(\sum_{j=1}^r m^{(j)}y_j\right)=\OO_Z\left(\sum_{x_i\in Z}\alpha_ix_i\right) 
\end{equation}

 If $Z$ is reducible, the correct extension is not to impose the same condition componentwise; rather, we should ask the numerical condition for the total degree, and, if $Z$ is obtained by gluing a smooth elliptic curve $E$ with a forest of rational trees $T_i$ at roots $q^E_i$, then the line bundle equality \eqref{eqn:eq_in_Pic} should be required in $\Pic(E)$ after counting each $q^E_i$ towards the left hand side with multiplicity
 \[f_*[T_i]\cdot [Y]+\sum_{\text{external components attached to $T_i$}} m^{(j)}- \sum_{x_j\in T_i}\alpha_j.\]

 %That this is the right thing to do follows from the proof that Gathmann's relative space is the projection of Jun Li's moduli space of maps to the expanded pair after contracting the higher levels of the accordion.
\end{remark}

>>>>

\newcommand{\pM}{\pazocal{M}}
\newcommand{\TT}{\operatorname{T}}
\newcommand{\oM}{\overline{\mathcal{M}}}
\newcommand{\M}[4]{\overline{\mathcal{M}}_{#1,#2}(#3,#4)}
\newcommand{\Q}[4]{\mathcal{Q}_{#1,#2}(#3,#4)}
\newcommand{\Qe}[4]{\mathcal{Q}^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\Qt}[4]{\widetilde{\mathcal Q}_{#1,#2}(#3,#4)}
\newcommand{\QG}[4]{\mathcal{Q}G_{#1,#2}(#3,#4)}
\newcommand{\QGe}[4]{\mathcal{Q}G^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\D}[3]{\mathcal{D^Q}(#1,#2,#3)}
\newcommand{\E}[3]{\mathcal{E^Q}(#1,#2,#3)}
\newcommand{\PP}{\mathbb P}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\VZ}{\pazocal{V\!Z}}
\newcommand{\tVZc}[4]{\widetilde{\mathcal{V\!Z}}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZc}[4]{\mathcal{V\!Z}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZcLi}[4]{\mathcal{V\!Z}^{\rm{ctr, Li}}_{#1,#2}(#3,#4)}
\newcommand{\VZrel}[4]{\mathcal{V\!Z}^{\rm{rel}}_{#1,#2}(#3,#4)}
\newcommand{\stab}{\rm{stab}}
\newcommand{\stC}{C'}
\newcommand{\stf}{f'}
\newcommand{\stpi}{\pi'}
\newcommand{\sts}{s'}
\newcommand{\N}{\mathbb{N}}
\newcommand{\OO}{\mathcal{O}}
\renewcommand{\to}{\rightarrow}
\newcommand{\A}{\mathcal A}
\newcommand{\B}{\mathcal B}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\EE}{\mathbf{E}}
\renewcommand{\L}{\mathcal L}
\newcommand{\LL}{\mathbf{L}}
\newcommand{\MM}{\mathfrak M}
\newcommand{\Aaff}{\mathbb{A}}
\newcommand{\kfield}{\Bbbk}
\newcommand{\comp}{\chi}
\newcommand{\sst}{\sigma^{\operatorname{ss}}}
\newcommand{\Pic}{\operatorname{Pic}}
\newcommand{\Def}{\operatorname{Def}}
\newcommand{\Spec}{\operatorname{Spec}}
\newcommand{\Proj}{\operatorname{Proj}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Gm}{\mathbb{G}_{\text{m}}}
\newcommand{\virt}[1]{[#1]^{\operatorname{virt}}}
\newcommand{\vip}[1]{[#1]^{\operatorname{prod}}}
\newcommand{\Id}{\operatorname{Id}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\HH}{\operatorname{H}}
\newcommand{\Achow}{\operatorname{A}}
\newcommand{\pt}{\operatorname{pt}}
\newcommand{\bq}{\begin{equation}}
\newcommand{\eq}{\end{equation}}
\newcommand{\ba}{\begin{aligned}}
\newcommand{\ea}{\end{aligned}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bsm}{\left(\begin{smallmatrix}}
\newcommand{\esm}{\end{smallmatrix}\right)}                   
\newcommand{\bpm}{\begin{pmatrix}}
\newcommand{\epm}{\end{pmatrix}}
\newcommand{\barr}{\begin{displaymath}\begin{array}{cccc}}
\newcommand{\earr}{\end{array}\end{displaymath}}
\newcommand{\barrl}{\begin{displaymath}\begin{array}{lcl}}
\newcommand{\earrl}{\end{array}\end{displaymath}}
\newcommand{\barl}{\begin{displaymath}\begin{array}{l}}
\newcommand{\earl}{\end{array}\end{displaymath}}
\newcommand{\bxym}{ \begin{displaymath}\xymatrix }
\newcommand{\exym}{\end{displaymath}}
\newcommand{\bcd}{\begin{center}\begin{tikzcd}}
\newcommand{\ecd}{\end{tikzcd}\end{center}}
\newcommand{\R}{\operatorname{R}^{\bullet}}
\newcommand{\dvr}{\Delta}
%\newcommand{\sslash}{\mathbin{/\mkern-6mu/}}
\newcommand{\tr}{{\rm tr}}
\newcommand{\Isom}{\text{Isom}}
\newcommand{\pr}{\operatorname{pr}}
\newcommand{\ev}{\operatorname{ev}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\vdim}{\operatorname{vdim}}
\newcommand{\ildef}[1]{\emph{#1}}
\newcommand{\om}[1]{\mathcal{#1}}
\newcommand{\h}{\operatorname{h}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\RR}{\textbf{R}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\ovm}[1]{\overline{\mathcal{#1}}}
\newcommand{\ovt}[1]{\widetilde{\mathcal{#1}}}
\newcommand{\ov}[1]{\overline{#1}}

@@

\section{Introduction}

\textbf{Statement of the problem.} Contrary to the genus zero case, the moduli space of genus one maps to projective space - with or without markings - is far from smooth; indeed it has various boundary components of different dimensions, representing maps that contract a genus one curve and have all the degree supported on a number of rational tails. The many incarnations of relative moduli spaces also suffer of the same undesirable feature.

Since the work of Vakil--Zinger and Ranganathan--Santos-Parker--Wise, it has been clear that it is possible to identify a desingularisation of the main component by adding the extra data of a contraction of the source curve $\nu\colon C\to \bar{C}$ - where the latter is allowed to acquire a Smyth singularity - and requiring the stable map $f\colon C\to \PP^N$ to factor through $\nu$.

\textbf{Choice of relative space and desingularisation.} We focus on the space of logarithmic stable maps to $(\PP^N|H)$, following ACGS. We perform a log modification of this space as detailed below. For a log curve $C\to S=\Spec(k=\bar k)$, modify the dual graph of $C$ by replacing the minimal genus one subcurve (in case it is a circle of $\PP^1$) by a single vertex of genus one, called the \emph{core} and denoted by $\circ$, and define a piecewise linear function with values in $\overline{\pazocal M}_S$ on such a graph by setting \[\lambda(v)=\sum_{q\in[\circ,v]}\rho_q,\]
where the $\rho_q$ are the smoothing parameters of the nodes $q$ separating $v$ from the core. Such a function is related to the log canonical bundle of $C\to S$. When the map contracts a subcurve of genus one, we endow it with the extra data of a radius $\delta\in\overline{\pazocal M}_S$ subject to the following compatibility condition:

(*) \emph{the circle of radius $\delta$ around $\circ$ passes through $\geq1$ vertex of positive $f$-degree}.

\noindent Furthermore, we require all the values of $\lambda$ to be comparable with $\delta$, and among themselves whenever they are $\leq \delta$. This is called a \emph{centrally aligned} log structure and carries enough information to define a contraction $\nu\colon C\to \bar{C}$, possibly after a semistabilisation of $(C,f)$ - in fact even more. The space thus obtained, $\widetilde \VZ_{1,\alpha}(\PP^N|H,d)$, is a log modification of $\M{1}{\alpha}{\PP^N|H}{d}$.

The main component $\VZ_{1,\alpha}(\PP^N|H,d)$ is then identified by a double factorisation condition:
\begin{enumerate}
 \item If $f$ contracts a genus one subcurve, then $f$ is required to factor through the Smyth singularity $\nu\colon C\to \bar C$ determined by the contraction radius $\delta$ as above.
 \item If furthermore the core is contracted by the associated tropical map $\phi$, let $\delta_2$ be the minimal distance from $\circ$ to a vertex supporting a flag that escapes $\phi^{-1}(\phi(\circ))$; we require $f$ to factor through $\nu_2\colon C\to\bar C_2$.
\end{enumerate}
The main result is that
\begin{thm}
 $\VZ_{1,\alpha}(\PP^N|H,d)$ is (log) smooth.
\end{thm}

\textbf{Gathmann's recursion.} There is a morphism $\VZ_{1,\alpha}(\PP^N|H,d)\to \VZ_{1,n}(\PP^N,d)$, hitting Gathmann's relative space. We may therefore pullback Gathmann's line bundle and section, cutting out the locus where the $k$-th marking is tangent to $H$ to order $\alpha_k+1$. Because $\alpha$ was maximal ($\sum\alpha=d$) by assumption, this means that the curve has to break, and $x_k$ has to lie on an internal component - one which is entirely mapped into $H$. Here comes an interesting point: the combinatorics of such boundary loci is governed by tropical geometry, and it is otherwise very hard to districate the interaction between the relative condition and the exceptional loci of the Vakil--Zinger blow-up.
\begin{cor}
 $\VZ_{1,\alpha}(\PP^N|H,d)$ is smooth over its Artin fan, in particular codimension one logarithmic strata can be read off from the latter.\marginpar{why are we only interested in log strata?}
\end{cor}
The Artin fan is a tropical gadget. Its local structure is given by subdividing the ACGS minimal model according to the alignment. We are only interested in picking its rays. The upshot is that the combinatorics is slightly more involved than in the genus zero case: the alignment may force some teeth of the comb to break.
\begin{thm}
 \[(\alpha_k\psi_k+\ev_k*H)[\VZ_{1,\alpha}(\PP^N|H,d)]=[D_{\alpha,k}(\PP^N|H,d)],\]
 where the latter is a sum of broken comb loci indexed by rays of the tropical fan.
\end{thm}
Importantly, the broken comb loci admit a very explicit description in terms of tautological integrals on the underlying boundary of Gathmann's relative space.
\begin{thm}
 Up to a finite cover of the underlying boundary stratum - which is just a combinatorially-determined product of moduli spaces of genus zero or one curves and absolute or relative maps with lower numerical invariants - every component of $D_{\alpha,k}(\PP^N|H,d)$ can be described as the transverse intersection of two loci in a projective bundle, where:
 \begin{itemize}
  \item the latter parametrises the possible line bundle isomorphisms imposed by the alignment of the log structure;
  \item the first locus is a subbundle representing the residual isomorphisms after fixing the ones imposed by tropical continuity;
  \item the second locus is determined by the factorisation conditions.
 \end{itemize}
\end{thm}
The upshot is that we may then push the formula down to the Gathmann's space and we obtain multiplicities and tautological classes.
\begin{cor}
 \[(\alpha_k\psi_k+\ev_k*H)[\VZ^G_{1,\alpha}(\PP^N|H,d)]=[D^G_{\alpha,k}(\PP^N|H,d)],\]
 where the latter is a weighted sum of tautological classes on Gathmann's comb loci.
\end{cor}
Once we have this formula, the following extensions are classical:
\begin{itemize}
 \item A similar formula for raising the tangency holds in the non-maximal case. It can be proven by adding auxiliary markings of contact order $1$; forgetting them is then a $(d-\sum\alpha)!:1$ cover because of the density of the nice locus inside Gathmann's spaces.
 \item The formula holds more generally for any smooth projective target $X$ relative to a generic hyperplane section $Y=X\cap H\subseteq \PP^N$ by virtual pullback.
\end{itemize}
Finally the recursive structure of the boundary allows us to prove the following
\begin{thm}[In-principle quantum Lefschetz]
 The restricted reduced genus one invariants of $Y$ can be inductively deducted from the full descendant (reduced) genus zero and one Gromov-Witten theory of $X$.
\end{thm}
The computation is more delicate than its genus zero analogue because invariant with the same numerical data appear intertwined in the last steps of the recursion.

@@
